---
title: "covid 19"
author: "Daniel Jones"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   

$H_0$ There is no difference between the amount of men compared to women that die from covid 19.    
$H_a$ There is a difference between the amount of men compared to women that die from covid19.  

# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="Percent Bar Chart")
```

The percentage of females dying is less than males. The percentage of males with the virus is less than the females but the rate of males dying is more than females.

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
Females alive is 56 percent and males is 44 percent.  
Females dead is 36 percent and males is 64 percent.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The odds of dying if your a male is 44/1411  
The odds of dying if your a female is 25/1831
The odds ratio is (44/1411)/(25/1831)  
The odds of dying are 2.3 to 1  

## Conclusions  
The p value is 0.001 which is less than 0.05  
The odds ratio is 2.3 to 1   
We reject the null hypothesis, there is no difference between men and women with regard to dying from covid 19.   
 
